Using a Bayesian hierarchical model to infer the redshift density function from photometric redshift probability distribution functions

### 2017-02-21 research/paper/draft.tex
unclear why the paper is altered in this branch but it will be overwritten by a merge from issue/41/reboot_paper eventually anyway

\shorttitle{How to estimate the redshift density function}
\shortauthors{Malz and Hogg}

@@

\title{How to estimate the redshift density distribution}

\author{Alex Malz\altaffilmark{1} \& David W. Hogg\altaffilmark{1,2,3,4}}
\email{aimalz@nyu.edu}

\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics,
  New York University, 4 Washington Pl., room 424, New York, NY 10003, USA}
\altaffiltext{2}{Simons Center for Data Analysis, 160 Fifth Avenue, 7th floor, New York, NY 10010, USA}
\altaffiltext{3}{Center for Data Science, New York University, 726 Broadway, 7th floor, New York, NY 10003, USA}
\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}

\begin{abstract}
The redshift density function $n(z)$ is a crucial ingredient for weak lensing cosmology.  Spectroscopically confirmed redshifts for the dim and numerous galaxies observed by past, present, and future weak lensing surveys are inaccessible, making photometric redshifts (photo-$z$s) the next best alternative.  Due to their intrinsic scatter and susceptibility to catastrophic outliers, photo-$z$ point estimates are being superseded by photo-$z$ probability distribution functions (PDFs).  However, analytic methods for utilizing these new data products in cosmological inference are still evolving.

This paper presents a novel approach to estimating the posterior distribution over $n(z)$ from a survey of galaxy photo-$z$ PDFs based upon a probabilistic graphical model.  We present the Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) code implementing this technique, as well as its validation on mock data and testing on a subset of BOSS DR10.  CHIPPR yields a more comprehensive and accurate characterization of $n(z)$ than traditional procedures.  The code is easily extensible to other one-point statistics of redshift that are informative to galaxy evolution and large-scale structure.

\end{abstract}

@@

\section{Introduction}
\label{sec:intro}

The redshift density function $n(z)$ is necessary for calculating two-point statistics of galaxy properties used to determine the cosmological parameter values that inform our understanding of the evolution of large-scale structure in the universe \citep{masters_mapping_2015}.  Inaccurate estimates of $n(z)$ can significantly impact the constraining power of a galaxy survey \citep{hildebrandt_kids-450:_2016}, biasing recovery of the cosmological parameters.

Though the redshift density function has traditionally been determined from spectroscopically observed redshifts, modern galaxy surveys seek to obtain two-point statistics of redshift from samples of galaxies for which spectroscopic redshifts are unavailable, either due to their large numbers or their low brightnesses.  For decades \citep{baum_photoelectric_1962}, photometrically estimated redshifts (photo-$z$s) have been the leading alternative to spectroscopically observed redshifts, though they suffer from issues of precision in the form of an intrinsic scatter and accuracy in the form of catastrophic outliers.

These weaknesses are illuminated by a probabilistic interpretation of photo-$zs$; if these complicated uncertainties were expressed as a probability distribution function (PDF) over redshift, photo-$z$s could be replaced by photo-$z$ PDFs containing more information than a simple point estimate.

A more recent

how surveys get $n(z)$

objective: given photo-$z$ PDFs, seek $n(z)$ PDF

%\appendix{}

\begin{acknowledgements}
AIM thanks Mohammadjavad Vakili for insightful input on statistics, Geoffrey Ryan for assistance in debugging code, and Boris Leistedt for helpful comments provided in the preparation of this paper.
\end{acknowledgements}

\bibliographystyle{apj}
\bibliography{references}

\end{document}

>>>>

\shorttitle{One-point galaxy population statistics from photo-$z$ PDFs}
\shortauthors{Malz and Hogg}

@@

\title{Using a Bayesian hierarchical model to infer the redshift density function from photometric redshift probability distribution functions}

\author{A.I. Malz\altaffilmark{1} \& David W. Hogg\altaffilmark{1,2,3,4} %\& Boris Leistedt\altaffilmark{1,5}
}
\email{aimalz@nyu.edu}

\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics,
  New York University, 4 Washington Pl., room 424, New York, NY 10003, USA}
\altaffiltext{2}{Simons Center for Data Analysis, 160 Fifth Avenue, 7th floor, New York, NY 10010, USA}
\altaffiltext{3}{Center for Data Science, New York University, 726 Broadway, 7th floor, New York, NY 10003, USA}
\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
%\altaffiltext{5}{Department of Physics \& Astronomy, University College London, Gower Street, London, WC1E 6BT, UK}

\begin{abstract}
Ongoing and upcoming galaxy surveys aiming to constrain cosmological parameters will produce photometric redshift (photo-$z$) probability distribution functions (PDFs), which are, generally, posteriors under an interim prior.  It is desirable to use this highly informative data product to estimate redshift-dependent distribution functions such as the redshift density function, stellar mass function, or We present a mathematically rigorous approach to obtain the posterior distribution on a one-point statistic of redshift from a survey providing photo-$z$ PDFs and their interim prior by way of a probabilistic graphical model corresponding to a hierarchical Bayesian relation of photo-$z$ PDFs to a posterior distribution over the one-point statistic.  This fully consistent method is applied to the redshift density function $n(z)$ necessary for the calculation of cosmological parameters from weak lensing surveys, validated on synthetic data and tested on a small subset of BOSS DR10 data.  It is demonstrated that this method improves accuracy, both qualitatively and as quantified by the Kullback-Leibler Divergence, compared to popular alternatives (such as stacking of photo-$z$ PDFs and conversion of photo-$z$ PDFs to point estimators of redshift).  The method can be straightforwardly generalized to other one-point statistics of redshift including distribution functions over luminosity, mass, star formation rate, etc.  We publicly release the Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) code that provides a general framework for estimating the posterior distribution of a general one-point statistic of redshift.
\end{abstract}

@@

%\clearpage
\section{Introduction}
\label{sec:intro}
